---
title: "tidymodels"
author: "firefreezing"
date: "4/10/2021"
output: github_document
---

```{r setup, include=FALSE, warning = F, message = F, comment = F}
# global knitr options
knitr::opts_chunk$set(fig.align = "center", echo = T, warning = F, message = F, comment = F)
```

### Nex steps:

1. try the workflow: add_recipe, add_model, fine_tune, etc.
2. try both training-validate-test, and cv + test
3. set a workflow to screen many models
4. benchmarking (running time)

### Useful functions:

+   `prep()`: execute the transformations from the recipe on top of the data that is supplied
+   `bake()`: turn the testing data into a preprocessed testing data following the recipe 
+   `juice()`: turn the training data into a preprocessed training data following the recipe


```{r load_pkgs}
library(tidymodels)

# Helper packages
library(readr)       # for importing data
library(broom.mixed) # for converting bayesian models to tidy tibbles
library(dotwhisker)  # for visualizing regression results
library(vip)
```

## Data preparation

You can also embed plots, for example:

```{r load_data, echo=FALSE}
hotels <- 
  read_csv('https://tidymodels.org/start/case-study/hotels.csv') %>%
  mutate_if(is.character, as.factor) 

dim(hotels)
```

```{r}
glimpse(hotels)
```

### Data splitting & resampling

```{r}
set.seed(123)

splits <- initial_split(hotels, prop = .8, strata = "children")

hotel_train <- training(splits)
hotel_test <- testing(splits)
```

Training data and testing data have similar proportion by `children`, thanks to the `strata` argument. 

```{r}
hotel_train %>%
  count(children) %>%
  mutate(prop = n/sum(n))
```
```{r}
hotel_test %>%
  count(children) %>%
  mutate(prop = n/sum(n))
```
## Recipe: data preprocessing

```{r}
holidays <- c("AllSouls", "AshWednesday", "ChristmasEve", "Easter", 
              "ChristmasDay", "GoodFriday", "NewYearsDay", "PalmSunday")

lr_recipe <- 
  recipe(children ~ ., data = hotel_train) %>% 
  step_date(arrival_date) %>% 
  step_holiday(arrival_date, holidays = holidays) %>% 
  step_rm(arrival_date) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())

lr_recipe
```

**prep**

`prep` executes the transformations on top of the data. 

```{r}
lr_recipe_exe <- lr_recipe %>% prep
```

**bake**

```{r}
# "juice" the processed training data
hotel_train_clean <- juice(lr_recipe_exe)

# "bake" the processed testing data
hotel_test_clean <- lr_recipe_exe %>%
  bake(new_data = hotel_test)
```

## Fit the prediction model

```{r}
cores <- parallel::detectCores()
cores
```

It takes 18.9s to run the model on single core
It takes 5.9s to run the model on an 8 core machine using `num.threads = cores`
```{r}
start_time <- Sys.time()

hotel_ranger <- rand_forest(trees = 100, mode = "classification") %>%
  set_engine("ranger") %>%
  fit(children ~ ., data = hotel_train_clean)

end_time <- Sys.time()

end_time - start_time
```

## Model performance

```{r}
hotel_ranger %>%
  predict(hotel_test_clean) %>%
  bind_cols(hotel_test_clean) %>%
  metrics(truth = children, estimate = .pred_class)
```

## Reference

+   An R community blog for a gentle introduction: https://rviews.rstudio.com/2019/06/19/a-gentle-intro-to-tidymodels/
+   Tidymodels official tutorial: https://www.tidymodels.org/start/
+   Tidy Modeling with R: https://www.tmwr.org/