---
title: "**Measure Testing for [Measure Name]**"
author: "QMD Stat Team"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: "cosmo"
    code_folding: hide
---

```{r packages_functions, message = FALSE, warning=FALSE, echo = F}
# load packages
library(pacman)
source("pkg_list.R")

# load helper functions
source("0_helper_fcts.r")

# global knitr options
opts_chunk$set(fig.align = "center", echo = T, warning = F, message = F, comment = F)
```

### **Decision log**
Add items here to reflect main decision updates during the process

+   01/01/2018 - update the analytic data by adding new records
+   02/02/2018 - revise the exclusion rule
+   03/03/2018 - change measre period from 2013-2014 to 2013-2015
+   etc.

### **Status**
Document the main milestones of the measure testing work. 

+   preliminary testing (done)
+   code review (in progress)
+   QAed by measure lead (not yet)

## **Measure Testing Results** {.tabset .tabset-fade .tabset-pills}


```{r, child = "1_measr_dat_intro.Rmd", eval = T}

```


```{r, child = "2_state_analysis.Rmd", eval = T}

```


```{r, child = "3_subgroup_analysis.Rmd", eval = T}

```


```{r, child = "4_reliability.Rmd", eval = T}

```

```{r, child = "5_validity.Rmd", eval = T}

```