---
title: "UpSet Plot for Measure Component Illustration"
author: "Fei"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: yes
  html_notebook:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
---

```{r packages_functions, message = FALSE, warning=FALSE, echo = FALSE}
# Load packages
library(UpSetR)
library(readr)
library(tidyverse)
library(lubridate)
library(knitr)

# Global knitr options
opts_chunk$set(fig.align = "center", echo = T, warning = F)
```

Read the measure data

```{r}
dat_measr <- read.csv("./data/dat_measure.csv")
```

### **Query particular intersection**

```{r}
upset(dat_measr, sets = c("denom", "denom_excld", "numer", "numer_excld"),
      keep.order = T, order.by = "freq", mb.ratio = c(.6, .4),
      query.legend = "bottom",
      queries = list(
        list(query = intersects, params = list("numer", "denom"),
             color = "steelblue", active = T, query.name = "Numerator  "),
        list(query = intersects, params = list("numer", "denom", "numer_excld"),
             color = "orange", active = T, query.name = "Numerator Exclusion")
      ),
      text.scale = 1.2,
      mainbar.y.label = "Component Intersections",
      sets.x.label = "Encounters per Component")
```

### **Query particular subgroup**

The follow green bar illustrates the proportion of male patients within each intersection component: 

```{r}
upset(dat_measr, sets = c("denom", "denom_excld", "numer", "numer_excld"),
      keep.order = T, order.by = "freq", mb.ratio = c(.6, .4),
      query.legend = "bottom",
      queries = list(
        list(query = elements, params = list("sex", 1),
             color = "#00694E", active = T, query.name = "Male")
      ),
      text.scale = 1.2,
      mainbar.y.label = "Component Intersections",
      sets.x.label = "Encounters per Component")
```

### **Customized query on attributes**

The follow green bar illustrates the proportion of male patients who are 65 and older within each intersection component: 

```{r}
myfunc <- function(row, age_thresh, sex) {
    data <- (row["age"] >= age_thresh) & (row["sex"] == sex)
}

upset(dat_measr, sets = c("denom", "denom_excld", "numer", "numer_excld"),
      keep.order = T, order.by = "freq", mb.ratio = c(.6, .4),
      query.legend = "bottom",
      queries = list(
        list(query = myfunc, params = list(65, 1),
             color = "#E70033", active = T, 
             query.name = "Male who are 65 and older")
      ),
      text.scale = 1.2,
      mainbar.y.label = "Component Intersections",
      sets.x.label = "Encounters per Component")
```

Another way to do it using expression:
```{r}
upset(dat_measr, sets = c("denom", "denom_excld", "numer", "numer_excld"),
      keep.order = T, order.by = "freq", mb.ratio = c(.6, .4),
      query.legend = "bottom",
      queries = list(
        list(query = myfunc, params = list(65, 1),
             color = "#E70033", active = T, 
             query.name = "Male who are 65 and older")
      ),
      expression = "sex == 1 & age >= 65",
      text.scale = 1.2,
      mainbar.y.label = "Component Intersections",
      sets.x.label = "Encounters per Component")
```

### **Build-in attribute plots associate with UpSet plot**

In this set of visualizations, the UpSet plot is the `active plot` (or the sender, if using terminology in communication theory), and all other scatter plots and histograms are the `passive plots` (or the receivers). 

```{r}
upset(dat_measr, sets = c("denom", "denom_excld", "numer", "numer_excld"),
      keep.order = T, order.by = "freq", mb.ratio = c(.6, .4),
      query.legend = "bottom",
      queries = list(
        list(query = intersects, params = list("numer", "denom"),
             color = "steelblue", active = T, query.name = "Numerator  "),
        list(query = intersects, params = list("numer", "denom", "numer_excld"),
             color = "orange", active = T, query.name = "Numerator Exclusion")
      ),
      text.scale = 1.2,
      mainbar.y.label = "Component Intersections",
      sets.x.label = "Encounters per Component",
      attribute.plots = list(gridrows = 100,
                             plots = list(list(plot = histogram, x = "age", queries = T),
                                          list(plot = histogram, x = "los", queries = T),
                                          list(plot = scatter_plot, x = "age", 
                                               y = "sex", queries = T),
                                          list(plot = scatter_plot, x = "age", 
                                               y = "los", queries = T)),
                             ncols = 2))
```

### **Boxplot of attribute by intersection**

Add boxplot to show the distribution of age (`age`) and length of stay (`los`) across all intersections.

```{r}
upset(dat_measr, sets = c("denom", "denom_excld", "numer", "numer_excld"),
      keep.order = T, order.by = "freq", mb.ratio = c(.6, .4),
      text.scale = 1.2,
      mainbar.y.label = "Component Intersections",
      sets.x.label = "Encounters per Component",
      boxplot.summary = c("age", "los"))
```


