---
title: "Use RShiny in RMarkdown"
author: "Fei"
date: "`r lubridate::today()`"
output: 
  html_document:
    code_folding: hide
    theme: cosmo
runtime: shiny
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)

library(tidyverse)
library(shiny)
# library(shinyTree)
```


This rmardown file provides some notes on RShiny, by various examples. Specifically, it covers the following information:

+  nesting `rshiny` in `rmarkdown`
+  `reactivity` features

### **Summary**

+ In general RShiny development, one can develop all the heavylifting components in the `server` side via `render` functions and leave the `ui` side relatively clean by just listing the arranagement of the UI components using `xxxOutput` functions. 

+ When writing RShiny in RMarkdown, one can break the traditional `ui` and `server` components and just flow the content within the R chunk. My understanding is that the R chunk in rmarkdown will handle the ui feature (which is essentially taking care of the html piece even without actually writing html).

+ All rshiny widgets can be viewed as special cases of `observeEvent`s.

+ Delay reactions using `eventReactive` allows to make update for multiple features and then trigger the change in output once the reactive values in 1st argument change (e.g. a click on the "GO" button). 


### **References**

1. [RShiny cheatsheet](https://shiny.rstudio.com/images/shiny-cheatsheet.pdf), by RStudio


#### **Example 1**: simple interactive visualization 

One can simply put all the UI outputs under the `ui` session use `xxxOutput` functions, and put the meat in the corresponding `renderxxx` function.

```{r, echo = T}
## example 1 - create a simple interactive app, and then move the ui piece to the server component,
## using renderUI

ui <- fluidPage(
  uiOutput(outputId = "ui"),
  br(),
  plotOutput(outputId = "hist")
)

server <- function(input, output){
  output$ui <- renderUI({
     sliderInput(inputId = "n", label = "Sample size", min = 20, max = 300, value = 20)
  })
  
  output$hist <- renderPlot({
    ggplot() +
      geom_histogram(aes(rnorm(input$n))) +
      labs(title = "Histogram of Normal Distribution", x = "value", y = "n") +
      theme_classic()
  })
}

shinyApp(ui = ui, server = server)
```

#### **Example 2**: ObserveEvent

`observeEvent()` runs code in 2nd argument when reactive values in 1st argument change. Hence, the above example can be viewed as an example under the `observeEvent` scenario, where a value in the sliderbar is the 1st argument, and the histogram (as 2nd argument) is updated whenever a new value is selected through the slidingbar. 

Here is a rewrite of the previous example:

```{r, echo = T}
# rewrite the previous email into
ui <- fluidPage(
  sliderInput(inputId = "n", label = "Sample size", min = 20, max = 300, value = 20),
  br(),
  plotOutput(outputId = "hist")
)

server <- function(input, output){
  observeEvent(input$n, {output$hist <- renderPlot({
      ggplot() +
        geom_histogram(aes(rnorm(input$n))) +
        labs(title = "Histogram of Normal Distribution", x = "value", y = "n") +
        theme_classic()
      })
  })
}

shinyApp(ui = ui, server = server)
```

Organize the code in a way such that all the feature settings are conducted (through `render`) in the `server` section.

```{r, echo = T}

ui <- fluidPage(
  uiOutput(outputId = "n"),
  br(),
  plotOutput(outputId = "hist")
)

server <- function(input, output){
  observeEvent(
    output$n <- renderUI({
      sliderInput(inputId = "n", label = "Sample size", min = 20, max = 300, value = 20)
      }), 
    {output$hist <- renderPlot({
      ggplot() +
        geom_histogram(aes(rnorm(input$n))) +
        labs(title = "Histogram of Normal Distribution", x = "value", y = "n") +
        theme_classic()
      })
  })
}

shinyApp(ui = ui, server = server)
```





#### **Example 3**: eventReactive

```{r, echo = T}

ui <- fluidPage(
  sliderInput(inputId = "n", label = "Sample size", min = 20, max = 300, value = 20),
  selectInput(inputId = "color", label = "Select bar color:",
              c("Blue" = "steelblue",
                "Red" = "red",
                "Green" = "green")),
  actionButton(inputId = "go", "Update"),
  plotOutput(outputId = "hist")
)

server <- function(input, output){
  plot_n <- eventReactive(
    input$go, {input$n})
  
  plot_color <- eventReactive(
    input$go, {input$color})
  
  output$hist <- renderPlot({
      ggplot() +
        geom_histogram(aes(rnorm(plot_n())), fill = plot_color()) +
        labs(title = "Histogram of Normal Distribution", x = "value", y = "n") +
        theme_classic()
      })
}

shinyApp(ui = ui, server = server)
```

Now re-organize the code flow slighly to make it a better `rshiny` and `rmarkdown` marriage. 

```{r}
sliderInput(inputId = "n", label = "Sample size", min = 20, max = 300, value = 20)

selectInput(inputId = "color", label = "Select bar color:",
            c("Blue" = "steelblue",
              "Red" = "red",
              "Green" = "green"))

actionButton(inputId = "go", "Update")

plot_n <- eventReactive(
  input$go, {input$n})

plot_color <- eventReactive(
  input$go, {input$color})
  
renderPlot({
  ggplot() +
    geom_histogram(aes(rnorm(plot_n())), fill = plot_color()) +
    labs(title = "Histogram of Normal Distribution", x = "value", y = "n") +
    theme_classic()
  })
```


#### **Backup materials**

```{r, eval = F, echo = F}
ui <- fluidPage(
  uiOutput(outputId = "ui"),
  
  actionButton("update1", "Update Here 1st"),
  
  uiOutput(outputId = "ui2")
  # ,
  # plotOutput(outputId = "hist")
)

server <- function(input, output){
  
  roster <- 1:5
  
  output$ui <- renderUI({
    checkboxGroupInput(inputId = "select1"
                       , label = "Select 1st subset to include for Analysis"
                       , choices = roster
                       , selected = NULL
                       , inline = TRUE)
  })

  
  
  dat <- data.frame(var1 = c(1,2,3,4,5), 
                    var2 = c("test1","test2","test3","test4","test5"))

  roster2 <- eventReactive(input$update1, dat$var2[dat$var1 %in% input$select1])
  # roster2 <- reactive(dat$var2[dat$var1 %in% input$select1])

  output$ui2 <- renderUI({
    checkboxGroupInput(inputId = "select2"
                   , label = "Select 2nd subset to include for Analysis"
                   , choices = roster2()
                   , selected = NULL
                   , inline = TRUE)
    })
}

shinyApp(ui = ui, server = server)
```


```{r, echo = F, eval = F}
roster <- 1:5

checkboxGroupInput(inputId = "select1"
                   , label = "Select 1st subset to include for Analysis"
                   , choices = roster
                   , selected = NULL
                   , inline = TRUE)

actionButton("update1", "Update Here 1st") 

dat <- data.frame(var1 = c(1,2,3,4,5), 
                  var2 = c("test1","test2","test3","test4","test5"))

roster2 <- eventReactive(input$update1, dat$var2[dat$var1 %in% input$select1])
# roster2 <- reactive(dat$var2[dat$var1 %in% input$select1])


# since using a reactive object here, need to render it in the "server" environment 
renderUI({
  checkboxGroupInput(inputId = "select2"
                   , label = "Select 2nd subset to include for Analysis"
                   , choices = roster2()
                   , selected = NULL
                   , inline = TRUE)
  })
```



